Hourly Assimilation of Different Sources of Observations Including Satellite Radiances in a  Mesoscale Convective System Case During RELAMPAGO. Part I: Impact on the Analysis and First-guess

### 2022-03-10 analysis/chapters/paper_ana_20181122.Rmd
Fix figures and captions

scale_date <- function(ini = 20181120180000, end = 20181123120000,
                       break_bin = "6 hour") {
  scale_x_datetime(breaks = seq.POSIXt(ymd_hms(ini), ymd_hms(end), by = break_bin),
                   labels = function(x) {
                     fifelse(hour(x) == 0, format(x, "%H\n%b %d"), format(x, "%H"))
                   }) 
}

@@

experiments <- c("E4", "E5", "E6", "E7")
#Color blind friendly but E4 y E7 parecidos en Desaturated
colores_exp <- c(E2 = "#0077BB", E5 = "#88CCEE", E6 = "#EE7733", E8 = "#CC3311") 
# colores_exp <- c(E4 = "#225555", E5 = "#CCBB44", E6 = "#CC6677", E7 = "#882255")
satinfo <- fread(here("analysis", "data", "derived_data", "satinfo.txt")) %>% 
  setnames(c("!sensor/instr/sat", "chan"), c("sensor", "channel"))
derived_data <- "/home/paola.corrales/datosmunin3/EXP/derived_data/"

@@

The analyses are generated using the Local Ensemble Transform Kalman Filter (LETKF) implementation (V1.3, @hunt2007) part of the Gridpoint Statistical Interpolation analysis system (GSI V3.8; @hu2018). 
A rapid update cycle approach is implemented with hourly analysis and a centered assimilation window, meaning that all the observations within $\pm$ 30 minutes of the analysis time are assimilated. 
Observations are assimilated in a 4D approach by comparing them with the corresponding first guess state at 10-minute intervals. 
For radiance observations, the Community Radiative Transfer Model version 2.3 (CRTM; @han2006) is used as an observation operator to calculate model-simulated brightness temperatures.

@@

(ref:dominio) a) The horizontal domain used on the simulations (black box), the inner domain used for the experiment comparison (red box), the region shown in b) (light blue box) and the locations of Automatic Surface Weather Stations (ASWS, green squares) and Conventional Surface Weather Stations (CSWS, orange triangles). b) Locations of radiosonde launches during the experiment, green dots corresponds to radiosondes launched during IOP 7, orange triangles are radiosondes launched during IOP 8 and purple square are radiosondes launched outside the IOP missions. The topography in meters is also shown (shaded).

@@

```{r table-obs}
fread(here("analysis/data/derived_data/table1.csv")) %>% 
  # str()
  kbl(caption = "Characteristics of the assimilated observations: The code for each observation type and its source, the available variables, the observation error and the gross check thresholds used.",
      col.names = c("Code", "Platform", "Variable", "Error", "Gross check"),
      booktabs = TRUE,
      escape = FALSE) %>% 
  kable_classic_2(full_width = FALSE) %>% 
  kable_styling(font_size = 6,
                position = "center") %>% 
  column_spec(1, width = "3.5em") %>% 
  column_spec(2, width = "4.5em") %>% 
  column_spec(3, width = "5em") %>% 
  column_spec(4:5, width = "7em") %>% 
  collapse_rows(columns = 1:2, valign = "middle", latex_hline = "major") %>% 
  footnote(symbol = c("Observation error varied with height.",
                      "Observations over 600 hPa are rejected.",
                      "Observation error depends on the report type."),
           symbol_manual = c("*", "**", "+"))
```

@@

```{r table-rad}
fread(here("analysis", "data", "derived_data", "tabla_radianzas.csv")) %>% 
  .[, ":="(sensor = toupper(sensor),
           plataforma = toupper(plataforma),
           prop = paste0(prop, " \\", "%"))] %>% 
  .[order(.[,'sensor']), ] %>% 
  .[] %>% 
  kbl(booktabs = TRUE,
      escape = FALSE,
      col.names = linebreak(c("Sensor", "Platform", "Assimilated\nchannels", "Percentage\nover total")),
      caption = "List of the available sensors over several platforms, the number of accepted channels for the assimilation and the percentage of assimilated observations calculated over all radiance observations and all cycles.") %>% 
  kable_styling(font_size = 7) %>% 
  collapse_rows(1, valign = "top") %>% 
  kable_classic_2(full_width = TRUE)
```

(ref:obs-horizontal) Horizontal spatial distribution of the mean available observations per analysis cycle for the a) CONV, b) ASWS c) SATWND and d) RAD experiments calculate over 2.5$^{\circ}$ boxes.

@@

files <- Sys.glob("/home/paola.corrales/datosmunin3/EXP/E8/ANA/*/diagfiles/asim*ensmean")

@@

rad <- read_diag_rad(files, "E8") %>% 
  satinfo[., on = c("sensor", "channel")] %>% 
  .[, ":="(lon.box = cut_round(lon, breaks = seq(284, 309, 2.5)),
           lat.box = cut_round(lat, breaks = seq(-42, -17, 2.5)))]

files <- Sys.glob("/home/paola.corrales/datosmunin3/EXP/E8/ANA/*/diagfiles/asim*ensmean")

@@

rad_conv <- read_diag_conv(files, exp = "E8", member = "000")

@@

rbind(conv, aut, satwnd, rad_conv) %>%
  .[, .(pressure, usage.flag, rerr, exp, date, lon.box, lat.box)] %>% 
  rbind(., rad) %>% 
  .[usage.flag == 1 & rerr != 1.0e+10] %>% 
  .[, .(count_obs = .N), by = .(exp, lon.box, lat.box, date)] %>% 
  .[, .(obs_cycle = mean(count_obs, na.rm = TRUE)), by = .(exp, lon.box, lat.box)] %>% 
  ggplot(aes(ConvertLongitude(lon.box), lat.box)) +
  geom_raster(aes(fill = obs_cycle), alpha = 0.8) +
  scale_fill_viridis_c(option = "D", trans = scales::log10_trans(),
                       guide = guide_colorbar(barwidth = 20,
                                              barheight = 0.5, 
                                              title.position = "left", 
                                              title.vjust = 1,
                                              frame.colour = "black")) +
  geom_mapa() +
  facet_wrap(~exp, labeller = labeller(exp = c(E2 = "CONV", E5 = "ASWS", 
                                               E6 = "SATWND", E8 = "RAD"))) +
  tagger::tag_facets() +
  labs(fill = "Mean number of\nobs per cycle") +
  theme_minimal(base_size = 13) +
  theme(legend.position = "bottom",
        tagger.panel.tag.text = element_text(size = 14),
        plot.margin = unit(c(0, 0, 0, 0), "cm"))
```

@@

rbind(conv, aut, satwnd, rad_conv) %>%
  .[, .(pressure, usage.flag, rerr, exp, date, lon.box, lat.box)] %>% 
  rbind(., rad) %>% 
  .[usage.flag == 1 & rerr != 1.0e+10] %>% 
  .[, .(count_obs = .N), by = .(exp, date)] %>% 
  ggplot(aes(date, count_obs)) +
  geom_line(aes(color = exp), size = 0.2) +
  geom_point(aes(color = exp, shape = exp), size = 1) +
  scale_color_manual(values = colores_exp, 
                     labels = c(E2 = "CONV", E5 = "ASWS", 
                                E6 = "SATWND", E8 = "RAD")) +
  scale_shape_manual(values = c(15, 16, 17, 18), labels = c(E2 = "CONV", E5 = "ASWS", 
                                                            E6 = "SATWND", E8 = "RAD")) +
  scale_y_log10() +
  scale_date() +
  labs(y = "Number of obs per cycle", x = "Hour (UTC)",
       color = NULL, shape = NULL) +
  theme_minimal(base_size = 10) +
  theme(legend.position = "bottom", 
        plot.margin = unit(c(0, 0, 0, 0), "cm")) +

  rbind(conv, aut, satwnd, rad_conv) %>%
  .[, .(pressure, usage.flag, rerr, exp, date, lon.box, lat.box)] %>% 
  rbind(., rad) %>%  
  .[usage.flag == 1 & rerr != 1.0e+10] %>% 
  .[, ":="(lev.box = cut_round(pressure, breaks = seq(50, 1050, 50)))] %>% 
  .[, .(count_obs = .N), by = .(exp, lev.box, date)] %>% 
  .[, .(obs_cycle = mean(count_obs, na.rm = TRUE)), by = .(exp, lev.box)] %>% 
  ggplot(aes(lev.box, obs_cycle)) +
  geom_line(aes(color = exp), size = 0.2) +
  geom_point(aes(color = exp, shape = exp), size = 1) +
  scale_color_manual(values = colores_exp, 
                     labels = c(E2 = "CONV", E5 = "ASWS", 
                                E6 = "SATWND", E8 = "RAD")) +
  scale_shape_manual(values = c(15, 16, 17, 18), labels = c(E2 = "CONV", E5 = "ASWS", 
                                                            E6 = "SATWND", E8 = "RAD")) +
  # geom_tile(aes(fill = obs_cycle)) +
  # scale_fill_viridis_c(option = "D", trans = scales::log10_trans(),
  #                      guide = guide_colorbar(barheight = NULL)) +
  scale_y_log10() +
  scale_x_reverse() +
  coord_flip() +
  labs(color = NULL, shape = NULL,
       x = "Pressure level (hPa)", y = "Mean number of\nobs per cycle") +
  theme_minimal(base_size = 10) +
  theme(legend.position = "bottom",
        plot.margin = unit(c(0, 0, 0, 0), "cm")) +

@@

rbind(fread(here("analysis", "data", "derived_data", "rcrv_V_perfil.csv")), 
      fread(here("analysis", "data", "derived_data", "rcrv_t_perfil.csv"))) %>% 
  obs.type[., on = "type"] %>%   
  .[type != 130] %>%
  melt(measure.vars = c("mean.y", "sd.y")) %>% 
  ggplot(aes(error.level, value)) +
  geom_hline(yintercept = c(0, 1), color = "grey20") +
  geom_line(aes(linetype = variable, color = code)) +
  scale_linetype_manual(values = c("dashed", "solid"),
                        labels = c("mean RCRV", "sd RCRV")) +
  scale_color_brewer(palette = "Dark2") +
  guides(linetype = guide_legend(order = 1),
         color = guide_legend(order = 10)) +
  # scale_color_discrete() +
  scale_x_level(breaks = c(1000, 500, 250, 125, 50)) +
  coord_flip() +
  facet_wrap(~ var, labeller = labeller(var = c("t" = "Temperature",
                                                "v" = "Wind"))) +
  tag_facets() +
  labs(x = "Pressure (hPa)", y = NULL,
       linetype = NULL, color = "Upper air\nobservations") +
  guides(color = guide_legend(keyheight = 0.4, default.unit = "cm"),
         linetype = guide_legend(keyheight = 0.4, default.unit = "cm")) +
  theme_minimal(base_size = 8) +
  theme(legend.box = "vertical",
        legend.spacing.y = unit(0.1, 'mm'),
        legend.title = element_text(size = 6),
        tagger.panel.tag.text = element_text(size = 8)) +

@@

  purrr::map(Sys.glob("/home/paola.corrales/mesoda/analysis/data/derived_data/rcrv_*_perfil.csv")[1:6], function(f) {
    meta <- unglue::unglue(basename(f), "rcrv_{sensor}_perfil.csv")
    fread(f) %>%
      .[, sensor := meta[[1]][["sensor"]]]
  }) %>%
  rbindlist() %>% 
  separate(sensor, into = c("sensor", "plat"), sep = "_") %>% 
  setDT() %>% 
  melt(measure.vars = c("mean.y", "sd.y")) %>% 
  .[, .(value = mean(value, na.rm = TRUE),
        var = "Brightness\ntemperature"), by = .(sensor, level, variable)] %>% 
  ggplot(aes(level, value)) +
  geom_hline(yintercept = c(0, 1), color = "grey20") +
  geom_line(aes(color = sensor, linetype = variable)) +
  scale_linetype_manual(values = c("dashed", "solid"),
                        labels = c("mean RCRV", "sd RCRV")) +
  scale_color_discrete(labels = toupper) +
  coord_flip() +
  scale_x_level(breaks = c(1000, 500, 250, 125, 50)) +
  facet_wrap(~ var, scales = "free_x") +
  tag_facets(tag_pool = c("d")) +
  labs(x = NULL, y = NULL,
       linetype = NULL, color = "Radiances\nobservations") +
  guides(color = guide_legend(keyheight = 0.4, default.unit = "cm"),
         linetype = guide_legend(keyheight = 0.4, default.unit = "cm")) +
  theme_minimal(base_size = 8) +
  theme(legend.box = "vertical",
        legend.spacing.y = unit(0.1, 'mm'),
        legend.title = element_text(size = 6),
        tagger.panel.tag.text = element_text(size = 8)) +

@@

(ref:TQ-diff) Difference between experiments ASWS-CONV (a and d), SATWND-ASWS (b and e) and RAD-SATWND (c and f) for the spatially averaged vertical profiles of temperature (a, b, and c, in $K$) and specific humidity (d, e, and f in $gkg^{-1}$) calculated over the inner domain (red box in Figure \@ref(fig:dominio)a) for each analysis cycle.

@@

perfiles %>% 
  dcast(lev + date ~ exp, value.var = "T") %>% 
  .[, ":="(E5_E2 = E5 - E2,
           E6_E5 = E6 - E5,
           E8_E6 = E8 - E6)] %>% 
  melt(measure.vars = c("E5_E2", "E6_E5", "E8_E6")) %>% 
  ggplot(aes(date, lev)) +
  geom_contour_fill(aes(z = value, fill = stat(level)),
                    breaks = c(seq(-2, 2, 0.2), Inf)) +
  geom_contour2(aes(z = value), color = "white", size = 0.1,
                breaks = c(seq(-2, 2, 0.2), Inf)) +
  scale_fill_divergent_discretised(guide = guide_colorsteps(barwidth = 25,
                                                            barheight = 0.3,
                                                            title.position = "left", 
                                                            title.vjust = 1,
                                                            frame.colour = "black")) +
  labs(fill = "Temperature (K)",
       x = NULL,
       y = NULL) +
  scale_y_level(breaks = c(1000, 850, 750, 500, 300, 200, 100)) +
  scale_date(ini = 20181121000000, break_bin = "12 hours") +
  facet_wrap(~variable, ncol = 3,
             labeller = labeller(variable = c("E5_E2" = "ASWS - CONV",
                                              "E6_E5" = "SATWND - ASWS",
                                              "E8_E6" = "RAD - SATWND"))) +
  tag_facets(position = list(x = 0.1, y = 0.96)) +
  theme_minimal(base_size = 8) +
  theme(legend.position = "bottom",
        tagger.panel.tag.text = element_text(size = 8),
        panel.ontop = TRUE,
        panel.grid = element_line(linetype = 3)) +

  perfiles %>% 
  dcast(lev + date ~ exp, value.var = "QVAPOR") %>% 
  .[, ":="(E5_E2 = E5 - E2,
           E6_E5 = E6 - E5,
           E8_E6 = E8 - E6)] %>% 
  melt(measure.vars = c("E5_E2", "E6_E5", "E8_E6")) %>% 
  ggplot(aes(date, lev)) +
  geom_contour_fill(aes(z = value*1000, fill = stat(level)), 
                    breaks = c(seq(-1.6, 1.6, 0.2), Inf)) +
  geom_contour2(aes(z = value), color = "white", size = 0.1,
                breaks = c(seq(-1.6, 1.6, 0.2), Inf)) +
  scale_fill_divergent_discretised(limits = c(-1.6, 1.6),
                                   guide = guide_colorsteps(barwidth = 25,
                                                            barheight = 0.3, 
                                                            title.position = "left", 
                                                            title.vjust = 1,
                                                            frame.colour = "black")) +

  scale_y_level(breaks = c(1000, 850, 750, 500, 300, 200, 100)) +
  scale_date(ini = 20181121000000, break_bin = "12 hours") +
  labs(fill = latex2exp::TeX("Specific \nhumidity ($gKg^{-1}$)"),
       y = NULL,
       x = NULL) +
  facet_wrap(~variable, ncol = 3,
             labeller = labeller(variable = c("E5_E2" = "ASWS - CONV",
                                              "E6_E5" = "SATWND - ASWS",
                                              "E8_E6" = "RAD - SATWND"))) +
  tag_facets(tag_pool = c("d", "e", "f"), position = list(x = 0.1, y = 0.96)) +
  theme_minimal(base_size = 8) +
  theme(legend.position = "bottom",
        tagger.panel.tag.text = element_text(size = 8),
        panel.ontop = TRUE,
        panel.grid = element_line(linetype = 3)) +

@@

(ref:UV-diff) Difference between experiments ASWS-CONV (a and d), SATWND-ASWS (b and e) and RAD-SATWND (c and f) for the spatially averaged vertical profiles of u wind (a, b, and c, in $ms^{-1}$) and v wind (d, e, and f in $ms^{-1}$) calculated over the inner domain (red box in Figure \@ref(fig:dominio)a) for each analysis cycle. In contours the u wind for ASWS (a), SATWND (b) and RAD (c) and v wind for ASWS (d), SATWND (e) and RAD (f) are shown.

>>>>

scale_date <- function(ini = 20181120180000, end = 20181123120000,
                       break_bin = "6 hour") {
  scale_x_datetime(breaks = seq.POSIXt(ymd_hms(ini), ymd_hms(end), by = break_bin),
                   labels = function(x) {
                     fifelse(hour(x) == 0, format(x, "%H UTC\n%b %d"), format(x, "%H"))
                   }) 
}

@@

experiments <- c("E4", "E5", "E6", "E7")
#Color blind friendly but E4 y E7 parecidos en Desaturated
colores_exp <- c(E2 = "#0077BB", E5 = "#88CCEE", E6 = "#EE7733", E9 = "#CC3311") 
# colores_exp <- c(E4 = "#225555", E5 = "#CCBB44", E6 = "#CC6677", E7 = "#882255")
satinfo <- fread(here("analysis", "data", "derived_data", "satinfo.txt")) %>% 
  setnames(c("!sensor/instr/sat", "chan"), c("sensor", "channel"))
derived_data <- "/home/paola.corrales/datosmunin3/EXP/derived_data/"

@@

The analyses are generated using the Local Ensemble Transform Kalman Filter (LETKF) implementation (V1.3, @hunt2007) part of the Gridpoint Statistical Interpolation analysis system (GSI V3.8; @hu2018). 
A rapid update cycle approach is implemented with hourly analysis and a centered assimilation window, meaning that all the observations within $\pm$ 30 minutes of the analysis time are assimilated. 
Observations are assimilated in a 4D approach by comparing them with the corresponding first-guess state at 10-minute intervals. 
For radiance observations, the Community Radiative Transfer Model version 2.3 (CRTM; @han2006) is used as an observation operator to calculate model-simulated brightness temperatures.

@@

(ref:dominio) a) The horizontal domain used for the simulations (black box), the inner domain used for the experiment comparison (red box), the region shown in b) (light blue box) and the locations of Automatic Surface Weather Stations (ASWS, green squares) and Conventional Surface Weather Stations (CSWS, orange triangles). b) Locations of radiosonde launches during the experiment. Green dots corresponds to radiosondes launched during IOP 7, orange triangles are radiosondes launched during IOP 8, and purple squares are radiosondes launched outside the IOP missions. The topography in meters is also shown (shaded).

@@

```{r table-obs}
fread(here("analysis/data/derived_data/table1.csv")) %>% 
  # str()
  kbl(caption = "Characteristics of the assimilated observations: The code for each observation type and its source, the available variables, the observation error and the gross check thresholds used.",
      col.names = c("Code", "Platform", "Variable", "Error", "Gross check"),
      booktabs = TRUE,
      escape = FALSE) %>% 
  kable_classic_2(full_width = FALSE) %>% 
  kable_styling(font_size = 6,
                position = "center") %>% 
  column_spec(1, width = "3.5em") %>% 
  column_spec(2, width = "4.5em") %>% 
  column_spec(3, width = "5em") %>% 
  column_spec(4:5, width = "7em") %>% 
  collapse_rows(columns = 1:2, valign = "middle", latex_hline = "major") %>% 
  footnote(symbol = c("Observation error varied with height.",
                      "Observations above 600 hPa are rejected.",
                      "Observation error depends on the report type."),
           symbol_manual = c("*", "**", "+"))
```

@@

```{r table-rad}
fread(here("analysis", "data", "derived_data", "tabla_radianzas.csv")) %>% 
  .[, ":="(sensor = toupper(sensor),
           plataforma = toupper(plataforma),
           prop = paste0(prop, " \\", "%"))] %>% 
  .[order(.[,'sensor']), ] %>% 
  .[] %>% 
  kbl(booktabs = TRUE,
      escape = FALSE,
      col.names = linebreak(c("Sensor", "Platform", "Assimilated\nchannels", "Percentage\nover total")),
      caption = "List of the available sensors over several platforms, the number of accepted channels for the assimilation, and the percentage of assimilated observations calculated over all radiance observations and all cycles.") %>% 
  kable_styling(font_size = 7) %>% 
  collapse_rows(1, valign = "top") %>% 
  kable_classic_2(full_width = TRUE)
```

(ref:obs-horizontal) Horizontal spatial distribution of the mean available observations per analysis cycle for the a) CONV, b) ASWS c) SATWND, and d) RAD experiments calculated over 2.5$^{\circ}$ boxes.

@@

files <- Sys.glob("/home/paola.corrales/datosmunin3/EXP/E9/ANA/*/diagfiles/asim*ensmean")

@@

rad <- read_diag_rad(files, "E9") %>% 
  satinfo[., on = c("sensor", "channel")] %>% 
  .[, ":="(lon.box = cut_round(lon, breaks = seq(284, 309, 2.5)),
           lat.box = cut_round(lat, breaks = seq(-42, -17, 2.5)))]

files <- Sys.glob("/home/paola.corrales/datosmunin3/EXP/E9/ANA/*/diagfiles/asim*ensmean")

@@

rad_conv <- read_diag_conv(files, exp = "E9", member = "000")

@@

rbind(conv, aut, satwnd, rad_conv) %>%
  .[, .(pressure, usage.flag, rerr, exp, date, lon.box, lat.box)] %>% 
  rbind(., rad) %>% 
  .[usage.flag == 1 & rerr != 1.0e+10] %>% 
  .[, .(count_obs = .N), by = .(exp, lon.box, lat.box, date)] %>% 
  .[, .(obs_cycle = mean(count_obs, na.rm = TRUE)), by = .(exp, lon.box, lat.box)] %>% 
  ggplot(aes(ConvertLongitude(lon.box), lat.box)) +
  geom_raster(aes(fill = obs_cycle), alpha = 0.8) +
  scale_fill_viridis_c(option = "D", trans = scales::log10_trans(),
                       guide = guide_colorbar(barwidth = 20,
                                              barheight = 0.5, 
                                              title.position = "left", 
                                              title.vjust = 1,
                                              frame.colour = "black")) +
  geom_mapa() +
  facet_wrap(~exp, labeller = labeller(exp = c(E2 = "CONV", E5 = "ASWS", 
                                               E6 = "SATWND", E9 = "RAD"))) +
  tagger::tag_facets() +
  labs(fill = "Mean number of\nobs per cycle") +
  theme_minimal(base_size = 13) +
  theme(legend.position = "bottom",
        tagger.panel.tag.text = element_text(size = 14),
        plot.margin = unit(c(0, 0, 0, 0), "cm"))
```

@@

rbind(conv, aut, satwnd, rad_conv) %>%
  .[, .(pressure, usage.flag, rerr, exp, date, lon.box, lat.box)] %>% 
  rbind(., rad) %>% 
  .[usage.flag == 1 & rerr != 1.0e+10] %>% 
  .[, .(count_obs = .N), by = .(exp, date)] %>% 
  ggplot(aes(date, count_obs)) +
  geom_line(aes(color = exp), size = 0.2) +
  geom_point(aes(color = exp, shape = exp), size = 1) +
  scale_color_manual(values = colores_exp, 
                     labels = c(E2 = "CONV", E5 = "ASWS", 
                                E6 = "SATWND", E9 = "RAD")) +
  scale_shape_manual(values = c(15, 16, 17, 18), labels = c(E2 = "CONV", E5 = "ASWS", 
                                                            E6 = "SATWND", E9 = "RAD")) +
  scale_y_log10() +
  scale_date() +
  labs(y = "Number of obs per cycle", x = "Hour (UTC)",
       color = NULL, shape = NULL) +
  theme_minimal(base_size = 10) +
  theme(legend.position = "bottom", 
        plot.margin = unit(c(0, 0, 0, 0), "cm")) +

  rbind(conv, aut, satwnd, rad_conv) %>%
  .[, .(pressure, usage.flag, rerr, exp, date, lon.box, lat.box)] %>% 
  rbind(., rad) %>%  
  .[usage.flag == 1 & rerr != 1.0e+10] %>% 
  .[, ":="(lev.box = cut_round(pressure, breaks = seq(50, 1050, 50)))] %>% 
  .[, .(count_obs = .N), by = .(exp, lev.box, date)] %>% 
  .[, .(obs_cycle = mean(count_obs, na.rm = TRUE)), by = .(exp, lev.box)] %>% 
  ggplot(aes(lev.box, obs_cycle)) +
  geom_line(aes(color = exp), size = 0.2) +
  geom_point(aes(color = exp, shape = exp), size = 1) +
  scale_color_manual(values = colores_exp, 
                     labels = c(E2 = "CONV", E5 = "ASWS", 
                                E6 = "SATWND", E9 = "RAD")) +
  scale_shape_manual(values = c(15, 16, 17, 18), labels = c(E2 = "CONV", E5 = "ASWS", 
                                                            E6 = "SATWND", E9 = "RAD")) +
  # geom_tile(aes(fill = obs_cycle)) +
  # scale_fill_viridis_c(option = "D", trans = scales::log10_trans(),
  #                      guide = guide_colorbar(barheight = NULL)) +
  scale_y_log10() +
  scale_x_reverse() +
  coord_flip() +
  labs(color = NULL, shape = NULL,
       x = "Pressure level (hPa)", y = "Mean number of\nobs per cycle") +
  theme_minimal(base_size = 10) +
  theme(legend.position = "bottom",
        plot.margin = unit(c(0, 0, 0, 0), "cm")) +

@@

rbind(fread(here("analysis", "data", "derived_data", "rcrv_V_perfil.csv")), 
      fread(here("analysis", "data", "derived_data", "rcrv_t_perfil.csv"))) %>% 
  obs.type[., on = "type"] %>%   
  .[type != 130] %>%
  melt(measure.vars = c("mean.y", "sd.y")) %>% 
  ggplot(aes(error.level, value)) +
  geom_hline(yintercept = c(0, 1), color = "grey20") +
  geom_line(aes(linetype = variable, color = code)) +
  scale_linetype_manual(values = c("dashed", "solid"),
                        labels = c("mean RCRV", "sd RCRV")) +
  scale_color_brewer(palette = "Dark2") +
  guides(linetype = guide_legend(order = 1),
         color = guide_legend(order = 10)) +
  # scale_color_discrete() +
  scale_x_level(name =  "Pressure (hPa)", breaks = c(1000, 500, 250, 125, 50)) +
  coord_flip() +
  facet_wrap(~ var, labeller = labeller(var = c("t" = "Temperature",
                                                "v" = "Wind"))) +
  tag_facets() +
  labs(x = "Pressure (hPa)", y = NULL,
       linetype = NULL, color = "Upper air\nobservations") +
  guides(color = guide_legend(keyheight = 0.4, default.unit = "cm"),
         linetype = guide_legend(keyheight = 0.4, default.unit = "cm")) +
  theme_minimal(base_size = 8) +
  theme(legend.box = "vertical",
        # legend.spacing.y = unit(0.1, 'mm'),
        legend.title = element_text(size = 6),
        tagger.panel.tag.text = element_text(size = 8)) +

@@

  purrr::map(Sys.glob("/home/paola.corrales/mesoda/analysis/data/derived_data/rcrv_*_perfil.csv")[1:6], function(f) {
    meta <- unglue::unglue(basename(f), "rcrv_{sensor}_perfil.csv")
    fread(f) %>%
      .[, sensor := meta[[1]][["sensor"]]]
  }) %>%
  rbindlist() %>% 
  .[sensor %in% c("airs", "amsua", "iasi", "mhs")] %>% 
  separate(sensor, into = c("sensor", "plat"), sep = "_") %>% 
  setDT() %>% 
  melt(measure.vars = c("mean.y", "sd.y")) %>% 
  .[, .(value = mean(value, na.rm = TRUE),
        var = "Brightness\ntemperature"), by = .(sensor, level, variable)] %>% 
  ggplot(aes(level, value)) +
  geom_hline(yintercept = c(0, 1), color = "grey20") +
  geom_line(aes(color = sensor, linetype = variable)) +
  scale_linetype_manual(values = c("dashed", "solid"),
                        labels = c("mean RCRV", "sd RCRV")) +
  scale_color_discrete(labels = toupper) +
  coord_flip() +
  scale_x_level(breaks = c(1000, 500, 250, 125, 50)) +
  facet_wrap(~ var, scales = "free_x") +
  tag_facets(tag_pool = c("d")) +
  labs(x = NULL, y = NULL,
       linetype = NULL, color = "Radiances\nobservations") +
  guides(color = guide_legend(keyheight = 0.4, default.unit = "cm"),
         linetype = guide_legend(keyheight = 0.4, default.unit = "cm")) +
  theme_minimal(base_size = 8) +
  theme(legend.box = "vertical",
        legend.spacing.y = unit(0.1, 'mm'),
        legend.title = element_text(size = 6),
        tagger.panel.tag.text = element_text(size = 8)) +

@@

(ref:TQ-diff) Difference between analysis ensemble mean experiments a) and b) ASWS-CONV, b) and e) SATWND-ASWS and c) and f) RAD-SATWND for the spatially averaged vertical profiles of temperature (a, b, and c, in $K$) and specific humidity (d, e, and f in $gkg^{-1}$) calculated over the inner domain (red box in Figure \@ref(fig:dominio)a) for each analysis cycle.

@@

perfiles %>% 
  dcast(lev + date ~ exp, value.var = "T") %>% 
  .[, ":="(E5_E2 = E5 - E2,
           E6_E5 = E6 - E5,
           E9_E6 = E9 - E6)] %>% 
  melt(measure.vars = c("E5_E2", "E6_E5", "E9_E6")) %>% 
  ggplot(aes(date, lev)) +
  geom_contour_fill(aes(z = value, fill = stat(level)),
                    breaks = c(seq(-2, 2, 0.2), Inf)) +
  geom_contour2(aes(z = value), color = "white", size = 0.1,
                breaks = c(seq(-2, 2, 0.2), Inf)) +
  scale_fill_divergent_discretised(guide = guide_colorsteps(barwidth = 25,
                                                            barheight = 0.3,
                                                            title.position = "left", 
                                                            title.vjust = 1,
                                                            frame.colour = "black")) +
  labs(fill = "Temperature (K)",
       x = NULL,
       y = "Pressure (hPa)") +
  scale_y_level(name = "Pressure (hPa)", breaks = c(1000, 850, 750, 500, 300, 200, 100)) +
  scale_date(ini = 20181121000000, break_bin = "12 hours") +
  facet_wrap(~variable, ncol = 3,
             labeller = labeller(variable = c("E5_E2" = "ASWS - CONV",
                                              "E6_E5" = "SATWND - ASWS",
                                              "E9_E6" = "RAD - SATWND"))) +
  tag_facets(position = list(x = 0.1, y = 0.96)) +
  theme_minimal(base_size = 8) +
  theme(legend.position = "bottom",
        tagger.panel.tag.text = element_text(size = 8),
        panel.ontop = TRUE,
        panel.grid = element_line(linetype = 3)) +

  perfiles %>% 
  dcast(lev + date ~ exp, value.var = "QVAPOR") %>% 
  .[, ":="(E5_E2 = E5 - E2,
           E6_E5 = E6 - E5,
           E9_E6 = E9 - E6)] %>% 
  melt(measure.vars = c("E5_E2", "E6_E5", "E9_E6")) %>% 
  ggplot(aes(date, lev)) +
  geom_contour_fill(aes(z = value*1000, fill = stat(level)), 
                    breaks = c(seq(-1.6, 1.6, 0.2), Inf)) +
  geom_contour2(aes(z = value), color = "white", size = 0.1,
                breaks = c(seq(-1.6, 1.6, 0.2), Inf)) +
  scale_fill_divergent_discretised(guide = guide_colorsteps(barwidth = 25,
                                                            barheight = 0.3, 
                                                            title.position = "left", 
                                                            title.vjust = 1,
                                                            frame.colour = "black")) +

  scale_y_level(name = "Pressure (hPa)", breaks = c(1000, 850, 750, 500, 300, 200, 100)) +
  scale_date(ini = 20181121000000, break_bin = "12 hours") +
  labs(fill = latex2exp::TeX("Specific \nhumidity ($gKg^{-1}$)"),
       y = "Pressure (hPa)",
       x = NULL) +
  facet_wrap(~variable, ncol = 3,
             labeller = labeller(variable = c("E5_E2" = "ASWS - CONV",
                                              "E6_E5" = "SATWND - ASWS",
                                              "E9_E6" = "RAD - SATWND"))) +
  tag_facets(tag_pool = c("d", "e", "f"), position = list(x = 0.1, y = 0.96)) +
  theme_minimal(base_size = 8) +
  theme(legend.position = "bottom",
        tagger.panel.tag.text = element_text(size = 8),
        panel.ontop = TRUE,
        panel.grid = element_line(linetype = 3)) +

@@

(ref:UV-diff)  Difference between analysis ensemble mean experiments a) and b) ASWS-CONV, b) and e) SATWND-ASWS and c) and f) RAD-SATWND for the spatially averaged vertical profiles of u wind (a, b, and c, in $ms^{-1}$) and v wind (d, e, and f in $ms^{-1}$) calculated over the inner domain (red box in Figure \@ref(fig:dominio)a) for each analysis cycle. Black contours corresponde to u wind for ASWS (a), SATWND (b) and RAD (c) and v wind for ASWS (d), SATWND (e) and RAD (f).

```{r UV-diff, fig.cap="(ref:UV-diff)", fig.env="figure*", fig.height=6, fig.align="center", fig.pos="ht"}
perfiles %>% 
  dcast(lev + date ~ exp, value.var = "U") %>% 
  .[, ":="(E5_E2 = E5 - E2,
           E6_E5 = E6 - E5,
           E9_E6 = E9 - E6)] %>% 
  melt(measure.vars = c("E5_E2", "E6_E5", "E9_E6")) %>% 
  ggplot(aes(date, lev)) +
  # geom_contour(aes(z = value)) +
  geom_contour_fill(aes(z = value, fill = stat(level)), breaks = c(seq(-2, 2, 0.2), Inf)) +
  scale_fill_divergent_discretised(guide = guide_colorsteps(barwidth = 25,
                                                            barheight = 0.3, 
                                                            title.position = "left", 
                                                            title.vjust = 1,
                                                            frame.colour = "black")) +
  geom_contour2(aes(z = value), color = "white", size = 0.1,
                breaks = c(seq(-2, 2, 0.2), Inf)) +
  geom_contour2(data = ~.x[, .(lev, date, E5_E2 = E5, E6_E5 = E6, E9_E6 = E9)] %>%
                  melt(id.vars = c("lev", "date")) %>% unique(), aes(z = value),
                breaks = seq(0, 30, 3), size = 0.2, linetype = 1, color = "grey30") +
  geom_text_contour(data = ~.x[, .(lev, date, E5_E2 = E5, E6_E5 = E6, E9_E6 = E9)] %>%
                      melt(id.vars = c("lev", "date")) %>% unique(), aes(z = value),
                    breaks = seq(0, 30, 3), color = "grey30", skip = 1,
                    size = 2, stroke = 0.1) +
  labs(fill = latex2exp::TeX("U ($ms^{-1}$)"),
       x = NULL,
       y = NULL) +
  scale_y_level(name = "Pressure (hPa)", breaks = c(1000, 850, 750, 500, 300, 200, 100)) +
  scale_date(ini = 20181121000000, break_bin = "12 hours") +
  facet_wrap(~variable, ncol = 3,
             labeller = labeller(variable = c("E5_E2" = "**ASWS** - CONV",
                                              "E6_E5" = "**SATWND** - ASWS",
                                              "E9_E6" = "**RAD** - SATWND"))) +
  tag_facets(position = list(x = 0.1, y = 0.96)) +
  theme_minimal(base_size = 8) +
  theme(legend.position = "bottom",
        tagger.panel.tag.text = element_text(size = 8),
        panel.ontop = TRUE,
        panel.grid = element_line(linetype = 3), 
        strip.text = ggtext::element_markdown()) +
